A User-Centric Approach to the Design and Consequences of Recommender Systems

### 2020-01-09 paper/filter_bubbles.tex
basic changes

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
We study a model of user decision-making in the context of recommender systems. We explore the extent to which two issues traditionally associated with recommender systems -- \textit{filter bubbles} and \textit{user homogenization} -- can be attributed to recommendation systems. The key component of our model is that users face a decision problem under uncertainty and that there are informational spillovers where a user consuming a product and learning how she values it reduces uncertainty about similar products. We find that this generates behavior consistent with filter bubble effects -- where users consume increasingly similar items over time -- and that recommendation can help alleviate these effects, as has been found in the empirical work of Nguyen et al. (2014). However, we find that recommendation, while alleviating filter bubble effects, leads to an increase in user homogeneity.
\par
We discuss the evaluation of recommender systems in light of these findings. Traditional design approaches to recommender systems only optimize for predicting user preferences after uncertainty has been resolved and ignore user beliefs and the inferences that users themselves make after consuming items. We propose an alternative interpretation of the serendipity measure where good recommendations are those that provide useful information, that is, information without which choice would be different and lead to worse outcomes on average relative to the case where no recommendation is provided. As a result, we emphasize the importance of understanding user behavior without recommendation, both for designing recommendations and understanding their impacts and usefulness.
\end{abstract}

@@

Recommender Systems (RS) have become critical for assisting users in navigating the large choice sets that they face in many online markets. For instance, users have to select from thousands of movies on Netflix, millions of products on Amazon, and billions of videos on YouTube. Users in many cases are not aware of most items, let alone have full information about their preferences over them and, to make matters worse, the goods in these markets are usually experience goods whose true value can only be learned after consumption. Furthermore, users interact with these systems routinely and watch more than one movie on Netflix, buy more than one product on Amazon, and listen to more than one video on YouTube.
\par

@@

\textbf{Our Model and Contributions} In this paper, 
%drawing from recent work in psychology and decision theory in economics, 
we first propose a model of consumer decision-making in the context of large choice sets faced by users in online markets and show that the model is able to rationalize features of user behavior in such contexts by means of simulations. 
We then ask how a stylized model of recommendation can shape the behavior of long-lived users
and finally ask how the insights from our model can be used to improve recommender system design and to better understand the consequences of these systems.
\par

>>>>

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
We study a model of user decision-making in the context of recommender systems. We explore the extent to which two issues -- \textit{filter bubbles} and \textit{user homogenization} -- can be attributed to recommendation systems. The key component of our model is that users face a decision problem under uncertainty and that there are informational spillovers where a user consuming a product and learning how she values it reduces uncertainty about similar products. We find that this generates behavior consistent with filter bubble effects -- where users consume increasingly similar items over time -- and that recommendation can help alleviate these effects, as has been found in the empirical work of Nguyen et al. (2014). However, we find that recommendation, while alleviating filter bubble effects, leads to an increase in user homogeneity.
\par
We discuss the evaluation of recommender systems in light of these findings. Traditional design approaches to recommender systems only optimize for predicting user preferences after uncertainty has been resolved and ignore user beliefs and the inferences that users themselves make after consuming items. We propose an alternative interpretation of the serendipity measure where good recommendations are those that provide useful information, without which choice would lead to worse outcomes on average relative to when no recommendation is provided. As a result, we emphasize the importance of understanding user behavior without recommendation, both for designing recommendations and understanding their impacts and usefulness.
\end{abstract}

@@

Recommender Systems (RS) have become critical for assisting users in navigating the large choice sets that they face in many online markets. For instance, users have to select from thousands of movies on Netflix, millions of products on Amazon, and billions of videos on YouTube. Users often are not aware of most items, let alone have full information about their preferences over them. To make matters worse, the goods in these markets are usually experience goods whose true value can only be learned after consumption. Furthermore, users interact with these systems routinely and watch more than one movie on Netflix, buy more than one product on Amazon, and listen to more than one video on YouTube.
\par

@@

\textbf{Our Model and Contributions} In this paper, 
%drawing from recent work in psychology and decision theory in economics, 
we propose a model of consumer decision-making in the context of large choice sets faced by users in online markets and show that the model is able to rationalize features of user behavior in such contexts by means of simulations. 
We then ask how a stylized model of recommendation can shape the behavior of long-lived users
and finally ask how the insights from our model can be used to improve recommender system design and to better understand the consequences of these systems.
\par
